---
title: "Exam 1"
format: html
---

```{r}
# part 1 Get the Data

food_consumption <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2020/2020-02-18/food_consumption.csv')

# Or read in with tidytuesdayR package (https://github.com/dslc-io/tidytuesdayR)
# PLEASE NOTE TO USE 2020 DATA YOU NEED TO USE tidytuesdayR version ? from GitHub

# Either ISO-8601 date or year/week works!

# Install via pak::pak("dslc-io/tidytuesdayR")


```
- Grand Research Question : What does the consumption of each food category in each country look like? 

- part 2 - put the code for installing packages in the console to avoid it being constantly redownloaded

```{r}
#part 3 

library(tidytuesdayR)
library(tidyverse)



```
- There are 9 packages loaded in teh tidyverse package 

```{r}
#part 4 

tuesdata <- tidytuesdayR::tt_load('2020-02-18')
tuesdata <- tidytuesdayR::tt_load(2020, week = 8)


fc <- tuesdata$food_consumption
```
- The code above gets data from tidytuesday and creates a new data set called fc specifically from food consumption and stores it there

```{r}
#part 5 

dim(fc) #gives dimension of data set 
head(fc) # loads first 6 rows 
tail(fc) #loads last 6 rows
str(fc) # gives structure of data 
summarise(fc) # gives a summary of the data
```
```{r}
#part 6 
head(fc,22)
tail(fc, 22)

fc |> group_by(food_category) |>
  summarise(total= n())
```
- The units of observation are country + food category . There are 11 food categories for 130 countries



- Part 7 
The grand research question involves 3 variables : country by food , food category and consumption 

```{r}
#Part 7

fc |>
  select(country, food_category, consumption)



```


- The grand research question involves 3 variables : country, food category and consumption 



```{r}
#part 9  - What does the consumption of each food category in each country look like? 

ggplot(fc, aes(x = country, y = consumption, fill = food_category)) +
  geom_bar(stat = "identity", width = 0.8) +
  labs(title = "Food Consumption by Category in Each Country",
       x = "Country",
       y = "Consumption (kg per capita)",
       fill = "Food Category") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) # Rotate country names for readability



```

